---
title: 'Task 2: Palmetto'
author: "Bri Baker"
date: "2/1/2021"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(GGally)
library(broom)
library(kableExtra)
```

## Introduction

Herein I utilize binary logistic regression to test the feasibility of using observations about individuals to differentiate between two palmetto species: *Serenoa repens* and *Sabal etonia*.

The variables utilized are plant height, canopy length, canopy width, and number of green leaves, and each annual observation is treated as an individual. Data were collected at Archbold Biological Station in south-central Florida from 1981 - 2017 at 5-year intervals (Abrahamson, 2019).


```{r data}

#could use col_types() if were using treatment

palmetto <- read_csv(here("task_2/palmetto.csv")) %>% # import data
  mutate(species_name = if_else(species == 1, "Serenoa repens", "Sabal etonia"), # add species names
         species_name = factor(species_name)) %>% # make names into factor for blm
  select(species, species_name, height:green_lvs) # select relevant columns

```


## Species Comparison

```{r green leaves, fig.cap= "**Figure 1.** Canopy length (cm) and green leaf count for *Sabal etonia* and *Serenoa repens*."}

# length vs leaves scatterplot

ggplot(data = palmetto, aes(x = length, y = green_lvs)) +
  geom_point(aes(color = species_name)) + 
  labs( x = "Canopy length (cm)",
        y = "Green Leaf Count",
        color = "Species",
        caption = "Bri Baker, 2021\nSource: Abrahamson, 2019") +
  scale_color_manual(labels = c("*Sabal etonia*" , "*Serenoa repens*"), # change legend names
                     values = c("goldenrod1", "cornflowerblue")) +
  theme_minimal() + # use theme_minimal
  theme(legend.text = ggtext::element_markdown(),
        legend.position = c(0.14, 0.86), # move legend
        legend.background = element_rect(fill="white", # format legend
                                         linetype = "solid", 
                                         color = "whitesmoke"))

```

Both species exhibit a positive correlation between canopy length and number of green leaves with *Serenoa repens* generally having more leaves at a given canopy length.


```{r height, fig.cap= "**Figure 2.** Comparison of canopy height (cm) of *Sabal etonia* and *Serenoa repens*."}

# boxplot of height by species

ggplot(data = palmetto, aes(x = species_name, y = height)) +
  geom_boxplot(aes(fill = species_name))+
  labs( x = "",
        y = "Canopy height (cm)",
        fill = "Species",
        caption = "Bri Baker, 2021\nSource: Abrahamson, 2019") +
  scale_fill_manual(labels = c("*Sabal etonia*" , "*Serenoa repens*"), # change legend names
                     values = c("goldenrod1", "cornflowerblue")) +
  theme_minimal() + # use theme_minimal
  theme(axis.text.x = element_text(face = "italic"),
        legend.position = "none") # remove legend

```

*Sabal etonia* and *Serenoa repens* are similar in canopy height although *Serenoa repens* is slightly larger on average with a wider spread in observed values. 


## Binary Logistic Regression

```{r}
palmetto_blr <- glm(species_name ~ height + length + width + green_lvs, # run binomial regression
                    data = palmetto,
                    family = "binomial")

palmetto_summary <- broom::tidy(palmetto_blr) %>% 
  mutate(p.value = case_when(p.value < 0.5 ~ "< 0.005"))# make table of outputs

kable(palmetto_summary, # make table
      caption = "Table 1: Summary statistics for palmetto species binomial logistic regression",
      col.names = c("Term", "Coefficient", "Standard error", "z value", "p value")) %>% 
  kable_paper(full_width = FALSE) # customize

```



## Model Success Assessment

```{r}
palmetto_fitted <- palmetto_blr %>% # use blr results
  broom::augment(type.predict = "response") %>% # make a table
  mutate(classified = case_when( # classify species based on results of blm
    .fitted < 0.5 ~ "Sabal etonia",
    .fitted >= 0.5 ~ "Serenoa repens"
  ),
  correct = case_when( # identify where species is correctly classified
    species_name == classified ~ "Yes",
    species_name != classified ~ "No"
  ))

summary_fit <- palmetto_fitted %>% # make summary graph
  count(species_name, correct) %>% # count correct/incorrect classification
  pivot_wider(names_from = correct, # switch organization
              values_from = n) %>% # with counts as values
  mutate(percent_correct = round(Yes/(No + Yes) * 100, 2)) # calculate % correct

kable(summary_fit, # make table
      caption = "Table 2: Summary of palmetto species binomial logistic regression classification success",
      col.names = c("Species", "Incorrect", "Correct", "Percent Correct (%)")) %>% 
  kable_paper(full_width = FALSE) # customize

```



## Citation

Data source: Abrahamson, W.G. 2019. *Survival, growth and biomass estimates of two dominant palmetto species of south-central Florida from 1981 - 2017, ongoing at 5-year intervals ver 1.* Environmental Data Initiative. https://doi.org/10.6073/pasta/f2f96ec76fbbd4b9db431c79a770c4d5